---
title: 2. Quantile Baseline
output: rmarkdown::html_document
---

This vignette describes the quantile baseline forecaster. The symmetrized
variation on this model is the `COVIDhub-baseline` model in the COVID-19
Forecast Hub.

We'll use these packages in this vignette:

```{r, message = FALSE}
library(covidcast)
library(epitools)
library(dplyr)
library(tidyr)
library(ggplot2)
library(simplets)
theme_set(theme_bw())
```

## Example data

We'll work with the following data with daily incident case counts of COVID-19 in Florida:

```{r fig.width=7, fig.height=7}
x <- covidcast_signal(data_source = "jhu-csse",
                      signal = "confirmed_incidence_num",
                      start_day = "2020-06-01",
                      end_day = "2021-05-31",
                      geo_type = "state",
                      geo_values = "fl",
                      as_of = "2021-10-28") %>%
  select(geo_value, time_value, cases = value) %>%
  as.epi_tibble()

head(x)
ggplot() +
  geom_line(data = x, mapping = aes(x = time_value, y = cases)) +
  geom_hline(yintercept = 0) +
  facet_wrap(vars(geo_value), scales = "free_y", ncol = 1)
```

Although the data presented here are *daily* data, we will illustrate a workflow
for generating one through four *week* ahead forecasts of weekly incident cases
in these locations using the quantile baseline model in combination with some
outlier correction and detection functionality from the `epitools` package.

# Outlier detection and correction

First, we do some cleanup, outlier detection and correction.  We iterate the
detection and correction procedure twice because nearby outliers are sometimes
hard to detect (the method thinks it is seeing a trend or high-variance period,
rather than nearby outliers).

```{r}
# drop trailing zeros
tail(x)
# x <- x[seq_len(nrow(x) - 3), ]


detection_methods = bind_rows(
  tibble(
    method = c("rolling_median"),
    method_args = list(list(detect_negatives = TRUE,
                            detection_multiplier = 2.5)),
    method_abbr = c("median")
  ),
  tibble(
    method = c("stl"),
    method_args = list(list(detect_negatives = TRUE,
                            detection_multiplier = 2.5)),
    method_abbr = c("stl_seasonal")
  ),
  tibble(
    method = c("stl"),
    method_args = list(list(detect_negatives = TRUE,
                            detection_multiplier = 2.5,
                            include_seasonality = FALSE)),
    method_abbr = c("stl_nonseasonal")
  )
)

x = x %>%
  # group_by(geo_value) to do outlier detection/correction separately for each
  # geo_value
  dplyr::group_by(geo_value) %>%
  detect_outliers(
    var = cases,
    methods = detection_methods,
    combine_method = "median",
    new_col_name = "outlier_info") %>%
  correct_outliers(
    var = cases,
    outliers_col = "outlier_info",
    detection_method = "combined",
    new_col_name = "corrected_cases1") %>%
  detect_outliers(
    var = corrected_cases1,
    methods = detection_methods,
    combine_method = "median",
    new_col_name = "outlier_info") %>%
  correct_outliers(
    var = corrected_cases1,
    outliers_col = "outlier_info",
    detection_method = "combined",
    new_col_name = "corrected_cases2")

head(x)
```

In the plot below, black is the original data, blue is after one round of
outlier detection and correction, and orange is after two rounds; we'll use the
data in orange as model inputs.

```{r}
ggplot(data = x) +
  geom_line(mapping = aes(x = time_value, y = cases)) +
  geom_line(mapping = aes(x = time_value, y = corrected_cases1), color = "cornflowerblue") +
  geom_line(mapping = aes(x = time_value, y = corrected_cases2), color = "orange") +
  theme_bw()
```


# Fitting baseline models

Here we fit a few variations on models:

```{r}
get_intervals_df <- function(predictions) {
  purrr::map_dfr(
    1:28,
    function(h) {
      purrr::map_dfr(
        c(0.025, 0.25),
        function(tau) {
          data.frame(
            h = h,
            level = as.character(100 * (1 - 2 * tau)),
            lower = quantile(predictions[, h], probs = tau),
            upper = quantile(predictions[, h], probs = 1 - tau)
          )
        }) %>%
        dplyr::mutate(level = factor(level, levels = c("95", "50")))
    })
}

get_baseline_predictions <- function(response_var,
                                     transformation,
                                     symmetrize,
                                     window_size) {
  # fit
  baseline_fit <- fit_simple_ts(
    y = x[[response_var]],
    ts_frequency = 1,
    model = 'quantile_baseline',
    transformation = transformation,
    transform_offset = ifelse(transformation == "none", 0, 1),
    d = 0,
    D = 0,
    symmetrize = symmetrize,
    window_size = window_size)

  # predict
  predictions <- predict(baseline_fit, nsim = 100000, horizon = 28, origin = "obs")

  # truncate to non-negative
  predictions <- pmax(predictions, 0)

  # extract predictive intervals and medians for plotting
  intervals_df <- get_intervals_df(predictions) %>%
    dplyr::mutate(t = nrow(x) + h)
  medians_df <- data.frame(
    t = nrow(x) + seq_len(ncol(predictions)),
    yhat = apply(predictions, 2, median)
  )

  return(tibble(
    intervals_df = list(intervals_df),
    medians_df = list(medians_df)
  ))
}

variations_to_fit <- tidyr::expand_grid(
  response_var = "corrected_cases2",
  transformation = c("none", "sqrt"),
  symmetrize = c(TRUE, FALSE),
  window_size = c(nrow(x) - 1, 28)
)

predictions <- purrr::pmap_dfr(variations_to_fit, get_baseline_predictions)
```

```{r}
intervals_df <- dplyr::bind_cols(variations_to_fit, predictions) %>%
  dplyr::select(-medians_df) %>%
  dplyr::mutate(model = paste("baseline",
                              "transform", transformation,
                              ifelse(symmetrize, "symmetrized", "non_symmetrized"),
                              "window_size", window_size,
                              sep = "_")) %>%
  tidyr::unnest(intervals_df)

medians_df <- dplyr::bind_cols(variations_to_fit, predictions) %>%
  dplyr::select(-intervals_df) %>%
  dplyr::mutate(model = paste("baseline",
                              "transform", transformation,
                              ifelse(symmetrize, "symmetrized", "non_symmetrized"),
                              "window_size", window_size,
                              sep = "_")) %>%
  tidyr::unnest(medians_df)
```

```{r, fig.width=10, fig.height=12}
observed_df <- dplyr::bind_rows(
  data.frame(
    t = seq_len(nrow(x)),
    y = x$cases,
    obs_type = "observed"
  ),
  data.frame(
    t = seq_len(nrow(x)),
    y = x$corrected_cases2,
    obs_type = "corrected"
  )
) %>%
  dplyr::mutate(obs_type = factor(obs_type, levels = c("observed", "corrected")))

p <- ggplot() +
  geom_line(mapping = aes(x = t, y = y, color = obs_type), data = observed_df) +
  geom_ribbon(mapping = aes(x = t, ymin = lower, ymax = upper, fill = level), data = intervals_df) +
  geom_line(mapping = aes(x = t, y = yhat), data = medians_df, color = "blue") +
  scale_color_manual(values = c("observed" = "orange", "corrected" = "black")) +
  scale_fill_viridis_d() +
  facet_wrap(vars(model), scales = "free_y", ncol = 2) +
  theme_bw()

print(p)
```
